Computing properties of independent sets by generic programming tensor networks \thanks{\funding{...}}

### 2022-01-08 paper/paper.tex
revision up to section 6.1

\centerline{\begin{tikzpicture}[]
    \scriptsize
    \node[ellipse,draw,minimum width=2cm, minimum height=1cm,inner sep=0pt] at (0, 2.5) (R) {$\mathbbm{R}$};
    % set and set sampler
    \node[ellipse,draw,minimum width=5cm, minimum height=3cm,inner sep=0pt] at (0, 0) (SN) {};
    \node[below of=1] at (SN) {Set};
    \node[ellipse,draw,minimum width=4cm, minimum height=1.5cm,inner sep=0pt] at (-0.5, 0) (S1) {\hspace{1.5cm}Bit string};
    \def\dx{-3};
    \node[ellipse,draw,minimum width=5cm, minimum height=3cm,inner sep=0pt] at (\dx, 0) (PN) {\hspace{-3.5cm}Polynomial};
    \node[ellipse,draw,minimum width=3.5cm, minimum height=1.2cm,inner sep=0pt] at (\dx, 0.7) (P1) {\hspace{-0.7cm}Highest order};
    \node[ellipse,draw,minimum width=3.5cm, minimum height=0.7cm,inner sep=0pt] at (\dx, -0.4) (P2) {\hspace{-0.7cm}2nd highest order};
    \node[ellipse,draw,minimum width=0.8cm, minimum height=0.3cm,inner sep=0pt] at (\dx, 1.0) (T) {};
    \node at (\dx, -1.2) {$\ldots$};
    \node[above of=2] at (T) (textT) {Tropical};
    \draw[black,-latex] (textT) -- (T);
\end{tikzpicture}}
\vspace{2em}
\centering{
\begin{minipage}{0.8\columnwidth}\scriptsize
\begin{tabularx}{\textwidth}{bb}
    \textbf{element type}     & \textbf{property to compute} \\
    {$\mathbbm{R}$}     & {the number of independent sets} \\
    {Polynomial} (\Eq{eq:polynomial})     & {independence polynomial} \\
    {Tropical (\Eq{eq:tropical})}    & {MIS size} \\
    {Polynomial truncated to highest order (\Eq{eq:countingtropical})}     & {MIS size and the number of MISs} \\
    {Polynomial truncated to 2nd highest order} (\Eq{eq:max2poly})     & {the number of MISs and independent sets of size $\alpha(G)-1$} \\
    {Set} (\Eq{eq:set})     & {enumeration of independent sets} \\
    {Bit string} (\Eq{eq:singleconfig})     & {one independent set} \\
    {Polynomial truncated to highest order combined with Bit string}     & {MIS size and one MIS} \\
    {Polynomial truncated to highest order combined with Set} (\Eq{eq:countingtropicalset})    & {MIS size and all MISs} \\
\end{tabularx}
\end{minipage}
}

@@

\section{Maximum independent sets and its counting}
\subsection{Tropical algebra for finding the MIS size and counting MISs}
In the previous section, we focused on computing the independence polynomial for a graph $G$ of given MIS size $\alpha(G)$, but we did not show how to compute this number.
The method we use to compute this quantity is based on the following observations. Let $x=\infty$, the independence polynomial becomes
\begin{equation}
I(G, \infty) = a_{\alpha(G)} \infty^{\alpha(G)},
\end{equation}
where the lower-order terms vanish. We can thus replace the polynomial type $a = (a_0, a_1, \ldots, a_k)$ with new type with two fields: the largest exponent $k$ and its coefficient $a_k$.
From this, we can define a new algebra as
%\begin{equation}
%\begin{split}
%    a(k) \oplus a(j) &= \begin{cases}
%        \left(a_k + a_j, \max(k,j) \right), & k = j \\
%        \left(a_j, \max(k,j) \right), & k < j \\
%        \left(a_k, \max(k,j) \right), & k > j
%    \end{cases}, \\
%    a(k) \odot a(j) &= (a_k a_j, k+j) \\
%    \mymathbb{0} &= (0, -\infty)\\
%    \mymathbb{1} &= (1, 0). \label{eq:countingtropical}
%\end{split}
%\end{equation}
\begin{equation}
    \eqname{P1}
\begin{split}
    a_x\infty^x \oplus a_y\infty^y &= \begin{cases}
        (a_x + a_y)\infty^{\max(x,y)}, & x = y\\
        a_y\infty^{\max(x,y)}, & x < y\\
        a_x\infty^{\max(x,y)}, & x > y
    \end{cases}, \\
    a_x\infty^x \odot a_y\infty^y &= a_x a_y\infty^{x+y}\\
    \mymathbb{0} &= 0\infty^{-\infty}\\
    \mymathbb{1} &= 1\infty^{0}.
\end{split}%& \text{\rlap{(P1)}}
\label{eq:countingtropical}
\end{equation}
To implement this algebra programmatically, we create a data type with two fields $(x, a_x)$ to store the MIS size and its counting,
and define the above operations and constants correspondingly.
When we are only interested in knowing the MIS size, we can drop the counting field.
The algebra of the exponents becomes the max-plus tropical algebra~\cite{Maclagan2015, Moore2011}.:
\begin{equation}\eqname{T}
    \begin{split}
        x \oplus y &= \max(x,y)\\
        x \odot y &= x + y\\
        \mymathbb{0} &= -\infty\\
        \mymathbb{1} &= 0.
    \end{split}\label{eq:tropical}
\end{equation}
This algebra is the same as the one used in Liu et al.~\cite{Liu2021} to calculate and count spin glass ground states.
For independent set calculations here, the vertex tensor and edge tensor becomes:
\begin{equation}
    W^{\rm tropical} = \left(\begin{matrix}
        \mymathbb{1} \\
        \infty
    \end{matrix}\right),   
    \qquad \qquad
        B^{\rm tropical} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}

\subsection{Truncated polynomial algebra for counting independent sets of large size}
Instead of counting just the MISs, one may be interested in counting the independent sets of large sizes close to the MIS size.
For example, if one is interested in counting only $a_{\alpha(G)}$ and $a_{\alpha(G)-1}$, we can define a truncated polynomial algebra by keeping only the largest two coefficients in the polynomial in \Eq{eq:polynomial}:
\begin{equation}
    \eqname{P2}
    \begin{split}
    a \oplus b &= (a_{\max(k_a, k_b)-1} + b_{\max(k_a, k_b)-1}, a_{\max(k_a, k_b)} + b_{\max(k_a, k_b)}),\\
    a \odot b &= (a_{k_a-1} b_{k_b}+a_{k_a} b_{k_b-1}, a_{k_a} b_{k_b}),\\
    \mymathbb{0} &= (), \\
    \mymathbb{1} &= (1).\label{eq:max2poly}
    \end{split}
\end{equation}
In the program, we thus need a data structure that contains three fields, the largest order $k$, and the coefficients for the two largest orders $a_k$ and $a_{k-1}$.
This approach can clearly be extended to calculate more independence polynomial coefficients and is more efficient than calculating the entire independence polynomial.
As will be shown below, this algebra can also be extended to enumerate those large-size independent sets.

\section{Enumeration of independent sets}
\subsection{Set algebra for configuration enumeration}
The enumeration problems of independent sets are also interesting and has been studied extensively in the literature~\cite{Bron1973, Eppstein2010, Johnson1988}, including,
for example, the enumeration of all independent sets, the enumeration of all maximal independent sets, or the enumeration of all MISs.
To enumerate all independent sets, we designed an algebra defined on sets of bitstrings.
\begin{equation}
\eqname{SN}
\begin{split}
    s \oplus t &= s \cup t\\
    s \odot t &= \{\sigma \lor^\circ \tau \, | \, \sigma \in s, \tau \in t\}\\
    \mymathbb{0} &= \{\}\\
    \mymathbb{1} &= \{0^{\otimes |V|}\}.
\end{split}\label{eq:set}
\end{equation}
where $s$ and $t$ are each a set of $|V|$-bit strings and $\lor^\circ$ is the bitwise OR operation over two bit strings.
\begin{example}\label{eg:setalgebra}
    For elements being bit strings of length $5$, we have the following set algebra
\begin{equation*}
\begin{split}
    &\{00001\} \oplus \{01110, 01000\} = \{01110, 01000\} \oplus \{00001\} = \{00001,01110, 01000\}\\
    &\{00001\} \oplus \{\} = \{00001\}\\
&\\
    &\{00001\} \odot \{01110, 01000\} = \{01110, 01000\} \odot \{00001\} = \{01111, 01001\}\\
    &\{00001\} \odot \{\} = \{\}\\
    &\{00001\} \odot \{00000\} = \{00001\}
\end{split}
\end{equation*}
\end{example}

To enumerate all independent sets, 
we initialize variable $x_{i}$ in the vertex tensor to $x_i = \{\boldsymbol{e}_{i}\}$, where $\boldsymbol{e}_i$ is a basis bit string of size $|V|$ that has only one non-zero value at location $i$.
The vertex and edge tensors are thus
\begin{equation}
    W^{\rm enum}(\{\boldsymbol{e}_{i}\}) = \left(\begin{matrix}
        \mymathbb{1} \\
        \{\boldsymbol{e}_{i}\}
    \end{matrix}\right),   
    \qquad \qquad
        B^{\rm enum} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}

This set algebra can serve as the coefficients in \Eq{eq:countingtropical} to enumerate all MISs, \Eq{eq:polynomial} to enumerate independent sets of different sizes,
or \Eq{eq:max2poly} to enumerate all independent sets of size $\alpha(G)$ and $\alpha(G)-1$.
As long as the coefficients in a truncated polynomial forms a commutative semiring, the polynomial itself is a commutative semiring.
For example, to enumerate only the MISs, with the tropical algebra, we define $s_{k}\infty^k$,
where the coefficient follows the algebra in \Eq{eq:set} and the orders follows the max-plus tropical algebra.
The combined operations become: 
%\begin{equation}
%\begin{split}
%    s(k) \oplus s(j) &= \begin{cases}
%        \left(s_k \cup s_j, \max(k,j) \right), & k = j \\
%        \left(s_j, \max(k,j) \right), & k < j \\
%        \left(s_k, \max(k,j) \right), & k > j
%    \end{cases}, \\
%    s(k) \odot s(j) &= (\{\sigma \lor^\circ \tau \, | \, \sigma \in s_k, \tau \in s_j\}, k+j) \\
%    \mymathbb{0} &= (\{ \}, -\infty)\\
%    \mymathbb{1} &= (\{0^{\otimes |V|}\}, 0). \label{eq:countingtropicalset}
%\end{split}
%\end{equation}
\begin{equation}
\eqname{P1+SN}
\begin{split}
    s_x\infty^x \oplus s_y\infty^y &= \begin{cases}
        (s_x \cup s_y)\infty^{\max(x,y)}, & x = y\\
        s_y\infty^{\max(x,y)}, & x < y\\
        s_x\infty^{\max(x,y)}, & x > y
    \end{cases},\\
    s_x\infty^x \odot s_y\infty^y &= \{\sigma \lor^\circ \tau | \sigma \in s_x, \tau \in s_y\}\infty^{x+y},\\
    \mymathbb{0} &= \{\}\infty^{-\infty},\\
    \mymathbb{1} &= \{0^{\otimes |V|}\}\infty^{0}. \label{eq:countingtropicalset}
\end{split}
\end{equation}
Clearly, the vertex tensor and edge tensor become
\begin{equation}
    W^{\rm MISenum}(\{\boldsymbol{e}_{i}\infty^1\}) = \left(\begin{matrix}
        \mymathbb{1} \\
        \{\boldsymbol{e}_{i}\}\infty^1
    \end{matrix}\right),   
    \qquad \qquad
        B^{\rm MISenum} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}
The contraction of the corresponding tensor network yields an enumeration of all MIS configurations.

@@

\subsection{Bounding the MIS enumeration space}
\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth, trim={0cm 0cm 0cm 0cm}, clip]{figures/masktree.pdf}
    \caption{Bounded enumeration of maximum independent sets. In the graph, a circle is a tensor, an arrow specifies execution direction of a function and $\circ$ is the Hadamard multiplication. (a) is the forward pass with algebra \Eq{eq:tropical} for computing $\alpha(G)$.
     (b) is the backward pass for computing boolean gradients as masks.
     (c) is the masked forward pass with algebra \Eq{eq:countingtropicalset} for enumerating configurations.}
     \label{fig:bounding}
\end{figure}

>>>>

\centerline{\begin{tikzpicture}[]
    \scriptsize
    \node[ellipse,draw,minimum width=2cm, minimum height=1cm,inner sep=0pt] at (0, 2.5) (R) {$\mathbbm{R}$};
    % set and set sampler
    \node[ellipse,draw,minimum width=5cm, minimum height=3cm,inner sep=0pt] at (0, 0) (SN) {};
    \node[below of=1] at (SN) {Set};
    \node[ellipse,draw,minimum width=4cm, minimum height=1.5cm,inner sep=0pt] at (-0.5, 0) (S1) {\hspace{1.5cm}Bit string};
    \def\dx{-3};
    \node[ellipse,draw,minimum width=5cm, minimum height=3cm,inner sep=0pt] at (\dx, 0) (PN) {\hspace{-3.5cm}Polynomial};
    \node[ellipse,draw,minimum width=3.5cm, minimum height=1.2cm,inner sep=0pt] at (\dx, 0.7) (P1) {\hspace{-0.7cm}Highest order};
    \node[ellipse,draw,minimum width=3.5cm, minimum height=0.7cm,inner sep=0pt] at (\dx, -0.4) (P2) {\hspace{-0.7cm}2nd highest order};
    \node[ellipse,draw,minimum width=0.8cm, minimum height=0.3cm,inner sep=0pt] at (\dx, 1.0) (T) {};
    \node at (\dx, -1.2) {$\ldots$};
    \node[above of=2] at (T) (textT) {Tropical};
    \draw[black,-latex] (textT) -- (T);
\end{tikzpicture}}
\vspace{2em}
{\centering{
\begin{minipage}{0.8\columnwidth}\scriptsize
\begin{tabularx}{\textwidth}{bb}
    \textbf{element type}     & \textbf{property to compute} \\
    {$\mathbbm{R}$}     & {the number of independent sets} \\
    {Polynomial} (\Eq{eq:polynomial})     & {independence polynomial} \\
    {Tropical (\Eq{eq:tropical})}    & {MIS size} \\
    {Polynomial truncated to highest order (\Eq{eq:countingtropical})}     & {MIS size and the number of MISs} \\
    {Polynomial truncated to 2nd highest order} (\Eq{eq:max2poly})     & {the number of MISs and independent sets of size $\alpha(G)-1$} \\
    {Set} (\Eq{eq:set})     & {enumeration of independent sets} \\
    {Bit string} (\Eq{eq:singleconfig})     & {one independent set} \\
    {Polynomial truncated to highest order combined with Bit string}     & {MIS size and one MIS} \\
    {Polynomial truncated to highest order combined with Set} (\Eq{eq:countingtropicalset})    & {MIS size and all MISs} \\
\end{tabularx}
\end{minipage}
}}

@@

\section{Maximum independent sets and its counting}
\subsection{Tropical algebra for finding the MIS size and counting MISs}
In the previous section, we focused on computing the independence polynomial for a graph $G$ of a given MIS size $\alpha(G)$, but we did not show how to compute this number.
The method we use to compute this quantity is based on the following observations. Let $x=\infty$, the independence polynomial becomes
\begin{equation}
I(G, \infty) = a_{\alpha(G)} \infty^{\alpha(G)},
\end{equation}
where the lower-order terms vanish. We can thus replace the polynomial type $a = (a_0, a_1, \ldots, a_k)$ with a new type with two fields: the largest exponent $k$ and its coefficient $a_k$.
From this, we can define a new algebra as
%\begin{equation}
%\begin{split}
%    a(k) \oplus a(j) &= \begin{cases}
%        \left(a_k + a_j, \max(k,j) \right), & k = j \\
%        \left(a_j, \max(k,j) \right), & k < j \\
%        \left(a_k, \max(k,j) \right), & k > j
%    \end{cases}, \\
%    a(k) \odot a(j) &= (a_k a_j, k+j) \\
%    \mymathbb{0} &= (0, -\infty)\\
%    \mymathbb{1} &= (1, 0). \label{eq:countingtropical}
%\end{split}
%\end{equation}
\begin{equation}
    \eqname{P1}
\begin{split}
    a_x\infty^x \oplus a_y\infty^y &= \begin{cases}
        (a_x + a_y)\infty^{\max(x,y)}, & x = y\\
        a_y\infty^{\max(x,y)}, & x < y\\
        a_x\infty^{\max(x,y)}, & x > y
    \end{cases}, \\
    a_x\infty^x \odot a_y\infty^y &= a_x a_y\infty^{x+y}\\
    \mymathbb{0} &= 0\infty^{-\infty}\\
    \mymathbb{1} &= 1\infty^{0}.
\end{split}%& \text{\rlap{(P1)}}
\label{eq:countingtropical}
\end{equation}
To implement this algebra programmatically, we create a data type with two fields $(x, a_x)$ to store the MIS size and its counting,
and define the above operations and constants correspondingly.
If one is only interested in finding the MIS size, one can drop the counting field.
The algebra of the exponents becomes the max-plus tropical algebra~\cite{Maclagan2015, Moore2011}:
\begin{equation}\eqname{T}
    \begin{split}
        x \oplus y &= \max(x,y)\\
        x \odot y &= x + y\\
        \mymathbb{0} &= -\infty\\
        \mymathbb{1} &= 0.
    \end{split}\label{eq:tropical}
\end{equation}
This algebra is the same as the one used in Liu et al.~\cite{Liu2021} to calculate and count spin glass ground states.
For independent set calculations here, the vertex tensor and edge tensor becomes:
\begin{equation}
    W^{\rm T} = \left(\begin{matrix}
        \mymathbb{1} \\
        \infty
    \end{matrix}\right),   
    \qquad \qquad
        B^{\rm T} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}

\subsection{Truncated polynomial algebra for counting independent sets of large size}
Instead of counting just the MISs, one may be interested in counting the independent sets of large sizes close to the MIS size.
For example, if one is interested in counting only $a_{\alpha(G)}$ and $a_{\alpha(G)-1}$, we can define a truncated polynomial algebra by keeping only the largest two coefficients in the polynomial in \Eq{eq:polynomial} as:
\begin{equation}
    \eqname{P2}
    \begin{split}
    a \oplus b &= (a_{\max(k_a, k_b)-1} + b_{\max(k_a, k_b)-1}, a_{\max(k_a, k_b)} + b_{\max(k_a, k_b)}),\\
    a \odot b &= (a_{k_a-1} b_{k_b}+a_{k_a} b_{k_b-1}, a_{k_a} b_{k_b}),\\
    \mymathbb{0} &= (), \\
    \mymathbb{1} &= (1).\label{eq:max2poly}
    \end{split}
\end{equation}
In the program, we thus need a data structure that contains three fields, the largest order $k$, and the coefficients for the two largest orders $a_k$ and $a_{k-1}$.
This approach can clearly be extended to calculate more independence polynomial coefficients and is more efficient than calculating the entire independence polynomial.
As will be shown below, this algebra can also be extended to enumerate those large-size independent sets.

\section{Enumeration of independent sets}
\subsection{Set algebra for configuration enumeration}
The enumeration problems of independent sets are also interesting and has been studied extensively in the literature~\cite{Bron1973, Eppstein2010, Johnson1988}, including,
for example, the enumeration of all independent sets, the enumeration of all maximal independent sets, or the enumeration of all MISs.
To enumerate all independent sets, we designed an algebra defined on sets of bitstrings:
\begin{equation}
\eqname{SN}
\begin{split}
    s \oplus t &= s \cup t\\
    s \odot t &= \{\sigma \lor^\circ \tau \, | \, \sigma \in s, \tau \in t\}\\
    \mymathbb{0} &= \{\}\\
    \mymathbb{1} &= \{0^{\otimes |V|}\}.
\end{split}\label{eq:set}
\end{equation}
where $s$ and $t$ are each a set of $|V|$-bit strings and $\lor^\circ$ is the bitwise OR operation over two bit strings.
\begin{example}\label{eg:setalgebra}
    For elements being bit strings of length $5$, we have the following set algebra
\begin{equation*}
\begin{split}
    &\{00001\} \oplus \{01110, 01000\} = \{01110, 01000\} \oplus \{00001\} = \{00001,01110, 01000\}\\
    &\{00001\} \oplus \{\} = \{00001\}\\
&\\
    &\{00001\} \odot \{01110, 01000\} = \{01110, 01000\} \odot \{00001\} = \{01111, 01001\}\\
    &\{00001\} \odot \{\} = \{\}\\
    &\{00001\} \odot \{00000\} = \{00001\}.
\end{split}
\end{equation*}
\end{example}

To enumerate all independent sets, 
we initialize variable $x_{i}$ in the vertex tensor to $x_i = \{\boldsymbol{e}_{i}\}$, where $\boldsymbol{e}_i$ is a basis bit string of size $|V|$ that has only one non-zero value at location $i$.
The vertex and edge tensors are thus
\begin{equation}
    W^{\rm SN}(\{\boldsymbol{e}_{i}\}) = \left(\begin{matrix}
        \mymathbb{1} \\
        \{\boldsymbol{e}_{i}\}
    \end{matrix}\right),   
    \qquad \qquad
        B^{\rm SN} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}

This set algebra can serve as the coefficients in \Eq{eq:polynomial} to enumerate independent sets of all different sizes, \Eq{eq:countingtropical} to enumerate all MISs,
or \Eq{eq:max2poly} to enumerate all independent sets of size $\alpha(G)$ and $\alpha(G)-1$.
As long as the coefficients in a truncated polynomial forms a commutative semiring, the polynomial itself is still a commutative semiring.
For example, to enumerate only the MISs, with the tropical algebra, we can define $s_{k}\infty^k$,
where the coefficients follow the algebra in \Eq{eq:set} and the orders (exponents) follow the max-plus tropical algebra.
The combined operations become: 
%\begin{equation}
%\begin{split}
%    s(k) \oplus s(j) &= \begin{cases}
%        \left(s_k \cup s_j, \max(k,j) \right), & k = j \\
%        \left(s_j, \max(k,j) \right), & k < j \\
%        \left(s_k, \max(k,j) \right), & k > j
%    \end{cases}, \\
%    s(k) \odot s(j) &= (\{\sigma \lor^\circ \tau \, | \, \sigma \in s_k, \tau \in s_j\}, k+j) \\
%    \mymathbb{0} &= (\{ \}, -\infty)\\
%    \mymathbb{1} &= (\{0^{\otimes |V|}\}, 0). \label{eq:countingtropicalset}
%\end{split}
%\end{equation}
\begin{equation}
\eqname{P1+SN}
\begin{split}
    s_x\infty^x \oplus s_y\infty^y &= \begin{cases}
        (s_x \cup s_y)\infty^{\max(x,y)}, & x = y\\
        s_y\infty^{\max(x,y)}, & x < y\\
        s_x\infty^{\max(x,y)}, & x > y
    \end{cases},\\
    s_x\infty^x \odot s_y\infty^y &= \{\sigma \lor^\circ \tau | \sigma \in s_x, \tau \in s_y\}\infty^{x+y},\\
    \mymathbb{0} &= \{\}\infty^{-\infty},\\
    \mymathbb{1} &= \{0^{\otimes |V|}\}\infty^{0}. \label{eq:countingtropicalset}
\end{split}
\end{equation}
Clearly, the vertex tensor and edge tensor become
\begin{equation}
    W^{\rm P1+SN}(\{\boldsymbol{e}_{i}\infty^1\}) = \left(\begin{matrix}
        \mymathbb{1} \\
        \{\boldsymbol{e}_{i}\}\infty^1
    \end{matrix}\right),   
    \qquad \qquad
        B^{\rm P1+SN} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}
The contraction of the corresponding tensor network yields an enumeration of all MIS configurations.

@@

\subsection{Bounding the MIS enumeration space}
\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth, trim={0cm 0cm 0cm 0cm}, clip]{figures/masktree.pdf}
    \caption{Bounded enumeration of maximum independent sets. In these graphs, a circle is a tensor, an arrow specifies execution direction of a function and $\circ$ is the Hadamard (element-wise) multiplication. (a) is the forward pass with algebra \Eq{eq:tropical} for computing $\alpha(G)$.
     (b) is the backward pass for computing boolean gradients as masks.
     (c) is the masked forward pass with algebra \Eq{eq:countingtropicalset} for enumerating configurations.}
     \label{fig:bounding}
\end{figure}
